---
title: "STAT 4355 Life Expectancy Project"
author: "Muhammad Munir, Vismaya Joseph, Caleb Captain"
date: "5/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Required Libraries

```{r}
shhh <- suppressPackageStartupMessages
shhh(library(ggplot2))
shhh(library(plotly))
shhh(library(tidyverse))
shhh(library(stringr))
shhh(library(mltools))
shhh(library(data.table))
shhh(library(car))
shhh(library(dplyr))
shhh(library(Hmisc))
shhh(library(corrplot))
shhh(library(MASS))
shhh(library(faraway))
shhh(library(viridis))
shhh(library(stargazer))
shhh(library(gtsummary))
shhh(library(sensemakr))
```

# Data Exploration and Cleaning

```{r}
data <- read.csv("Life Expectancy Data.csv")
describe(data)
```

These are the features of the data.
Also, the number of unique entries for each feature.
Some of these we will want to drop.
Country is a good example: 
we would not like the model to learn life expectancy from a country name.
It needs to learn from the values, we can use the names to as an outlier check.

```{r}
colnames(data)
unique(data$Status)
unique(data$Country)

Developed <- unique(data$Country[which(data$Status == "Developed")])
Developing <- unique(data$Country[which(data$Status == "Developing")])
```
```{r}
table(Developed)
```

```{r}
table(Developing)
```

This section of code shows us which features/columns have null values.

```{r}
colnames(data)[sapply(data, anyNA)]
table(is.na(data))
```

There are 2,563 null values in the data. 
But they are confined to 1,293 rows of null values.

```{r}
table(is.na.data.frame(data$Hepatitis.B))
table(is.na.data.frame(data$Life.expectancy))
table(is.na.data.frame(data$Polio))
table(is.na.data.frame(data$BMI))
table(is.na.data.frame(data$Diphtheria))
table(is.na.data.frame(data$thinness..1.19.years))
table(is.na.data.frame(data$thinness.5.9.years))
table(is.na.data.frame(data$Total.expenditure))
table(is.na.data.frame(data$Population))
table(is.na.data.frame(data$Income.composition.of.resources))
table(is.na.data.frame(data$Adult.Mortality))
table(is.na.data.frame(data$Schooling))
table(is.na.data.frame(data$Alcohol))
table(is.na.data.frame(data$GDP))
```

These are the features/columns with null values.

```{r}
colnames(data)[sapply(data, anyNA)]
```

```{r}
summary(data$Life.expectancy)
summary(data$Hepatitis.B)
summary(data$Total.expenditure)
summary(data$Population)
summary(data$Income.composition.of.resources)
summary(data$Adult.Mortality)
summary(data$BMI)
summary(data$Diphtheria)
summary(data$thinness..1.19.years)
summary(data$Schooling)
summary(data$Alcohol)
summary(data$Polio)
summary(data$GDP)
summary(data$thinness.5.9.years)
```

Boxplots and Histigrams of each of the null columns/features is shown below.

```{r}
ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Life.expectancy))) + 
           labs(x="Life Expectancy"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Hepatitis.B))) + 
           labs(x="Hepatitis B"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Total.expenditure))) + 
           labs(x="Total Expenditure"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Population))) + 
           labs(x="Population"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Income.composition.of.resources))) + 
           labs(x="Income Composition of Resources"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Adult.Mortality))) + 
           labs(x="Adult Mortality"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(BMI))) + 
           labs(x="Body-Mass Index"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Diphtheria))) + 
           labs(x="Diphtheria"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(thinness..1.19.years))) + 
           labs(x="Thinness of 10-19 Years"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Schooling))) + 
           labs(x="Schooling"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Alcohol))) + 
           labs(x="Alcohol"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(Polio))) + 
           labs(x="Polio"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(GDP))) + 
           labs(x="Gross Domestic Product"))

ggplotly(ggplot(data) + 
           geom_histogram(stat="count", aes(x=is.na(thinness.5.9.years))) + 
           labs(x="Thinness of 5-9 Years"))
```

Note that population is flat.
Essentially the population is so sparse that it does not show a range of values.

```{r}
ggplotly(ggplot(data, aes(y=Life.expectancy)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Hepatitis.B)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Total.expenditure)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Population)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Income.composition.of.resources)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Adult.Mortality)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=BMI)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Diphtheria)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=thinness..1.19.years)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Schooling)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Alcohol)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=Polio)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=GDP)) + geom_boxplot())
ggplotly(ggplot(data, aes(y=thinness.5.9.years)) + geom_boxplot())
```


```{r}
data <- as.data.table(data)

data1 <- data %>% filter(
  !is.na(data$Country),
  !is.na(data$Status),
  !is.na(data$Adult.Mortality),
  !is.na(data$Life.expectancy),
  !is.na(data$infant.deaths),
  !is.na(data$Alcohol),
  !is.na(data$percentage.expenditure),
  !is.na(data$Hepatitis.B),
  !is.na(data$Measles),
  !is.na(data$BMI),
  !is.na(data$under.five.deaths),
  !is.na(data$Polio),
  !is.na(data$Total.expenditure),
  !is.na(data$Diphtheria),
  !is.na(data$HIV.AIDS),
  !is.na(data$GDP),
  !is.na(data$Population),
  !is.na(data$thinness..1.19.years),
  !is.na(data$thinness.5.9.years),
  !is.na(data$Income.composition.of.resources),
  !is.na(data$Schooling)
)

data1$Status <- factor(x=data1$Status, levels=c("Developing", "Developed"), labels= c(0,1))
data1 <- one_hot(as.data.table(data1), cols = "Status")
data1$Status <- data1$Status_1
```

# Graphing

These graphs show the relations between Life Expectancy and various features of interest.

```{r}
ggplotly(ggplot(data1, aes(x=Alcohol, y=Life.expectancy)) +
           geom_point() +
           geom_count(aes(colour= ..n.., size=..n..)) +
           scale_color_viridis(option = "C") +
           theme(axis.title.x=element_text(angle=0,size = rel(1.25), 
           margin = margin(1, unit = "cm"),vjust = 1)) + 
           theme(axis.title.y=element_text(angle=90,size = rel(1.25))) +
           geom_smooth(method="lm") +
           labs(x="Alcohol Intake"))

ggplotly(ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           geom_point(alpha=.5, colour="blue") +
           geom_count(aes(colour= ..n.., size=..n..)) +
           scale_color_viridis(option="D") +
           theme(axis.title.x=element_text(angle=0,size = rel(1.25), 
           margin = margin(1, unit = "cm"),vjust = 1)) + 
           theme(axis.title.y=element_text(angle=0,size = rel(1.25))) + 
           labs(x="Development Status"))

ggplotly(ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           geom_boxplot() +
           theme(axis.title.x=element_text(angle=0,size = rel(1.25), 
           margin = margin(1, unit = "cm"),vjust = 1)) + 
           theme(axis.title.y=element_text(angle=90,size = rel(1.25))) +
           facet_wrap(data1$Status) +
           labs(x="Development Status"))
```


```{r}
ggplotly(ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           theme(axis.title.x=element_text(angle=0,size = rel(1))) + 
           theme(axis.title.y=element_text(angle=90, size = rel(1))) +
           geom_boxplot() + 
           facet_grid(rows = vars(data1$Status), cols=vars(data1$Year), scales = "free"))

ggplotly(ggplot(data1, aes(x=Schooling, y=Life.expectancy)) + 
           geom_point(alpha=.5, colour="blue") +
           geom_smooth(method="lm") + theme(axis.title.x=element_text(angle=0,size = rel(1.25), 
           margin = margin(1, unit = "cm"),vjust =1)) + 
           theme(axis.title.y=element_text(angle=0,size = rel(1.25))))
```


```{r}
mean(data1$Life.expectancy[which(data1$Year == 2000)])
mean(data1$Life.expectancy[which(data1$Year == 2015)])
```


```{r}
ggplotly(ggplot(data1, aes(x=Year, y=Life.expectancy)) + geom_point(alpha=.25) +
  geom_count(aes(colour= ..n.., size=..n..)) + 
  scale_color_viridis(option = "A") +
  geom_smooth(method="lm") + 
    theme(axis.title.x=element_text(angle=0,size = rel(1.5), 
    margin = margin(1, unit = "cm"),vjust =1)) +
    theme(axis.title.y=element_text(angle=0,size = rel(1.5))))
```


```{r}
ggplotly(ggplot(data1, aes(x= Income.composition.of.resources,y= Life.expectancy)) + 
  geom_point() +
  geom_count(aes(colour= ..n.., size=..n..)) +
  scale_color_viridis(option = "C") +
  geom_smooth(method="lm"))

ggplotly(ggplot(data1, aes(x= GDP,y= Life.expectancy)) + 
  geom_point())

ggplotly(ggplot(data1, aes(x= Total.expenditure,y= Life.expectancy)) + 
  geom_point())
```


```{r}
ggplotly(ggplot(data1, aes(x=Schooling,y=Life.expectancy)) +
           geom_count(aes(colour=..n..,size=..n..)) +
           facet_wrap(data1$Status) + 
           scale_color_viridis(option="D") +
           geom_smooth(method="lm"))

ggplotly(ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           geom_point(alpha=.5, colour="blue") +
           geom_count(aes(colour= ..n.., size=..n..)) +
           scale_color_viridis(option="D") +
           theme(axis.title.x=element_text(angle=0,size = rel(1.25), 
           margin = margin(1, unit = "cm"),vjust = 1)) + 
           theme(axis.title.y=element_text(angle=0,size = rel(1.25))))

ggplotly(ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           theme(axis.title.x=element_text(angle=0,size = rel(1))) + 
           theme(axis.title.y=element_text(angle=90, size = rel(1))) +
           geom_boxplot() + 
           facet_grid(rows = vars(data1$Status), cols=vars(data1$Year), scales = "free"))

ggplot(data1, aes(x=Status, y=Life.expectancy)) + 
           theme(axis.title.x=element_text(angle=0,size = rel(1))) + 
           theme(axis.title.y=element_text(angle=90, size = rel(1))) +
           geom_boxplot() + 
           facet_grid(rows = vars(data1$Status), cols=vars(data1$Year), scales = "free")
```

# Model Fitting

The first model for variable selection

```{r}
data1 <- within(data1, rm(Year, Country, Adult.Mortality, infant.deaths, Status_0, Status_1))

model1 <- lm(Life.expectancy ~ Status+Alcohol+percentage.expenditure+ 
               Hepatitis.B+Measles+BMI+under.five.deaths+Polio+Total.expenditure+ 
               Diphtheria+HIV.AIDS+GDP+Population+thinness..1.19.years+ 
               thinness.5.9.years+Income.composition.of.resources+Schooling, data1)

summary(model1)

hist(studres(model1), 
     breaks=10, freq=F, col="cornflowerblue", cex.axis=1.5, cex.lab=1.5, cex.main=2)

vif(model1)

bc1 <- boxCox(data1$Life.expectancy ~ data1$Status+data1$Alcohol+
         data1$percentage.expenditure+data1$Hepatitis.B+data1$Measles+
         data1$BMI+data1$under.five.deaths+data1$Polio+data1$Total.expenditure+
         data1$Diphtheria+data1$HIV.AIDS+data1$GDP+data1$Population+
         data1$thinness..1.19.years+data1$thinness.5.9.years+
         data1$Income.composition.of.resources+data1$Schooling)
```

Notice the histogram: It is an obvious normal distribution, so we won't need
transformations.

The second model is one with the optimized variables selected

```{r}
model2 <- lm(Life.expectancy ~ Status+Alcohol+percentage.expenditure+ 
               Hepatitis.B+Measles+BMI+under.five.deaths+Polio+Total.expenditure+ 
               Diphtheria+HIV.AIDS+Income.composition.of.resources+Schooling, data1)

summary(model2)
hist(studres(model2), 
     breaks=10, freq=F, col="cornflowerblue", cex.axis=1.5, cex.lab=1.5, cex.main=2)

bc2 <- boxCox(data1$Life.expectancy~data1$Status+data1$Alcohol+
         data1$percentage.expenditure+data1$Hepatitis.B+data1$Measles+
         data1$BMI+data1$under.five.deaths+data1$Polio+data1$Total.expenditure+
         data1$Diphtheria+data1$HIV.AIDS+
         data1$Income.composition.of.resources+data1$Schooling)
```

Another normal distribution. No transformation needed.

#Correlation Diagram

```{r}
typeof(data1)
data1.cor <- cor(data1)

corrplot(data1.cor, tl.cex=.8)
```

#Residual Analysis

```{r}
summary(residualPlot(model1, type="rstudent", quadratic=F, 
             col = "dodgerblue", pch=16, cex=1.5, 
             cex.axis=1.5, cex.lab=1.5, ylim=c(-3,3)))

residualPlot(model1, variable='fitted', type='rstudent')

qqPlot(model1)
```


```{r}
influenceIndexPlot(model1, vars=c('Cook'), data1, main = "Cook Diagnostic Plot")
influenceIndexPlot(model1, vars=c('hat'), data1, main = "Hat Diagnostic Plot")
influenceIndexPlot(model1, vars=c('Studentized'), data1, main = "Studentized Diagnostic Plot")
influencePlot(model1)
dfbetaPlots(model1, intercept=T)

influenceIndexPlot(model1, vars=c('Cook'), data1, main = "Cook Diagnostic Plot")
influenceIndexPlot(model1, vars=c('hat'), data1, main = "Hat Diagnostic Plot")
influenceIndexPlot(model1, vars=c('Studentized'), data1, main = "Studentized Diagnostic Plot")
influencePlot(model1)
dfbetaPlots(model1, intercept=T)

```

From the future: Outliers/Problem indexes are 903, 1122, 665, 672, 1344, 1339.

```{r}
data1[c(903,1122,665,672,1344,1339,355),]
```

Lets Take them out and try it again.

```{r}
data1 <- data1[-c(903,1122,665,672,1344,1339),]
data1

model1 <- lm(Life.expectancy ~ Status+Alcohol+percentage.expenditure+ 
               Hepatitis.B+Measles+BMI+under.five.deaths+Polio+Total.expenditure+ 
               Diphtheria+HIV.AIDS+GDP+Population+thinness..1.19.years+ 
               thinness.5.9.years+Income.composition.of.resources+Schooling, data1)

summary(model1)

hist(studres(model1), 
     breaks=10, freq=F, col="cornflowerblue", cex.axis=1.5, cex.lab=1.5, cex.main=2)

qqPlot(model1)
```

```{r}
influenceIndexPlot(model1, vars=c('Cook'), data1, main = "Cook Diagnostic Plot")
influenceIndexPlot(model1, vars=c('hat'), data1, main = "Hat Diagnostic Plot")
influenceIndexPlot(model1, vars=c('Studentized'), data1, main = "Studentized Diagnostic Plot")
influencePlot(model1)
dfbetaPlots(model1, intercept=T)
```

```{r}
summary(residualPlot(model2, type="rstudent", quadratic=F, 
             col = "dodgerblue", pch=16, cex=1.5, 
             cex.axis=1.5, cex.lab=1.5, ylim=c(-3,3)))

qqPlot(model2)

residualPlot(model2, variable='fitted', type='rstudent')
```

```{r}
influenceIndexPlot(model2, vars=c('Cook'), data1, main = "Cook Diagnostic Plot")
influenceIndexPlot(model2, vars=c('hat'), data1, main = "Hat Diagnostic Plot")
influenceIndexPlot(model2, vars=c('Studentized'), data1, main = "Studentized Diagnostic Plot")
influencePlot(model2)

dfbetaPlots(model2, intercept=T)

hist(studres(model2), 
     breaks=10, freq=F, col="cornflowerblue", cex.axis=1.5, cex.lab=1.5, cex.main=2)
```

Problem indices: 903, 355, 1122, 1344, 1339.
We took out 903, 1122, 1344, and 1339, so now we only should remove 355

```{r}
data1 <- data1[-c(355),]
data1

model2 <- lm(Life.expectancy ~ Status+Alcohol+percentage.expenditure+
               Hepatitis.B+Measles+BMI+under.five.deaths+Polio+Total.expenditure+
               Diphtheria+HIV.AIDS+Income.composition.of.resources+Schooling, data1)

summary(model2)
```

```{r}
influenceIndexPlot(model2, vars=c('Cook'), main = "Cook Diagnostic Plot")
influenceIndexPlot(model2, vars=c('hat'), main = "Hat Diagnostic Plot")
influenceIndexPlot(model2, vars=c('Studentized'), main = "Studentized Diagnostic Plot")
dfbetaPlots(model2, intercept=T)

hist(studres(model2), 
     breaks=10, freq=F, col="cornflowerblue", cex.axis=1.5, cex.lab=1.5, cex.main=2)

residualPlot(model2, variable='fitted', type='rstudent')

vif(model2)
```
