---
title: "Life Expectancy Analysis"
author: "VMC"
date: "4/27/2021"
output: html_document
---

```{r}
library(rmarkdown)
library (ggplot2)
life_expectancy_data <- read.csv("Life Expectancy Data.csv", header = TRUE)
```

# Scatter Plots
```{r}
ggplot(life_expectancy_data) + geom_hex(aes(Year, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))
life_expectancy_data$Developed <- as.integer (as.logical(life_expectancy_data$Status == "Developed"))

ggplot (life_expectancy_data) + geom_hex(aes(Developed, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Alcohol, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(percentage.expenditure, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Hepatitis.B, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Measles, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(BMI, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(under.five.deaths, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Polio, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Total.expenditure, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Diphtheria, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(HIV.AIDS, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(GDP, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Population, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(thinness..1.19.years, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(thinness.5.9.years, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Income.composition.of.resources, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Schooling, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

```

# Histograms
```{r}

hist (life_expectancy_data$Life.expectancy, xlab="Life Expectancy", ylab="Frequency of Life Expectancy", border="green",col="blue", main = "Life Expectancy")

```


# Fitting the model

In the following step, we will perform hypothesis testing to figure out variables that have a high linear corelation with life expectancy to arrive at a fitted multiple linear regression model

```{r}

render("life_expectancy_analysis.Rmd", "pdf_document")
```


# Residual Analysis

In the following step, we will perform residual analysis, to figure out outliers and variables that contribute for and against the accuracy of our model.



# Response to Feedback

How many data points do you have? Is each country a data point or a combination of country and year a data point? (-1)

Describe members' responsibility on the planned data analysis, not about the proposal. (-1)

Exploratory data analysis described in GOAL is better placed in PLAN.

# this is updated text


