---
title: "Life Expectancy Analysis"
author: "VMC"
date: "4/27/2021"
output: html_document
---

```{r}
library(rmarkdown)
library (ggplot2)
life_expectancy_data <- read.csv("Life Expectancy Data.csv", header = TRUE)
```

# Scatter Plots
```{r}
ggplot(life_expectancy_data) + geom_hex(aes(Year, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))
life_expectancy_data$Developed <- as.integer (as.logical(life_expectancy_data$Status == "Developed"))

ggplot (life_expectancy_data) + geom_hex(aes(Developed, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Alcohol, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(percentage.expenditure, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Hepatitis.B, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Measles, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(BMI, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(under.five.deaths, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Polio, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Total.expenditure, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Diphtheria, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(HIV.AIDS, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(GDP, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Population, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(thinness..1.19.years, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(thinness.5.9.years, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Income.composition.of.resources, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

ggplot (life_expectancy_data) + geom_hex(aes(Schooling, Life.expectancy), bins = 100) +
         scale_fill_gradientn("", colours = rev(rainbow(10, end = 4/6)))

```

# Histograms
```{r}

hist (life_expectancy_data$Life.expectancy, xlab="Life Expectancy", ylab="Frequency of Life Expectancy", border="green",col="blue", main = "Life Expectancy")

```


# Fitting the model

In the following step, we will perform hypothesis testing to figure out variables that have a high linear corelation with life expectancy to arrive at a fitted multiple linear regression model

```{r}

render("life_expectancy_analysis.Rmd", "pdf_document")
```


# Residual Analysis

In the following step, we will perform residual analysis, to figure out outliers and variables that contribute for and against the accuracy of our model.

```{r}


colnames(life_expectancy_data)
unique(life_expectancy_data$Status)
life_expectancy_data$Status <- factor(x=life_expectancy_data$Status, levels=unique(life_expectancy_data$Status), labels=c(0,1))
life_expectancy_data <- within(life_expectancy_data, rm(Country))
life_expectancy_data

life_expectancy_data$Life.expectancy[is.na(life_expectancy_data$Life.expectancy)] <- mean(life_expectancy_data$Life.expectancy, na.rm = T)
life_expectancy_data$Polio[is.na(life_expectancy_data$Polio)] <- mean(life_expectancy_data$Polio, na.rm = T)
life_expectancy_data$BMI[is.na(life_expectancy_data$BMI)] <- mean(life_expectancy_data$BMI, na.rm = T)
life_expectancy_data$Diphtheria[is.na(life_expectancy_data$Diphtheria)] <- mean(life_expectancy_data$Diphtheria, na.rm = T)
life_expectancy_data$Year[is.na(life_expectancy_data$Year)] <- mean(life_expectancy_data$Year, na.rm = T)
life_expectancy_data$percentage.expenditure[is.na(life_expectancy_data$percentage.expenditure)] <- mean(life_expectancy_data$percentage.expenditure, na.rm = T)
life_expectancy_data$HIV.AIDS[is.na(life_expectancy_data$HIV.AIDS)] <- mean(life_expectancy_data$HIV.AIDS, na.rm = T)
life_expectancy_data$Status[is.na(life_expectancy_data$Status)] <- mean(life_expectancy_data$Status, na.rm = T)



fit <- lm(life_expectancy_data$Life.expectancy~life_expectancy_data$Status+life_expectancy_data$percentage.expenditure+life_expectancy_data$BMI+life_expectancy_data$Polio+life_expectancy_data$Diphtheria+life_expectancy_data$HIV.AIDS)
summary(fit)

library(car)
library(MASS)
par(mfrow=c(1,2))
qqPlot(fit)

par(mfrow=c(1,1))
residualPlot(fit, type="rstudent", quadratic=F, col = "dodgerblue",
             pch=16, cex=1.5, cex.axis=1.5, cex.lab=1.5)

```

```{r}
# head (life_expectancy_data)
# plot(qqnorm (life_expectancy_data$percentage.expenditure)); qqline (life_expectancy_data$percentage.expenditure)
# life_expectancy_data$percentage.expenditure
# log.per <- log (life_expectancy_data$percentage.expenditure)
# log.per[log.per = -Inf] <- mean (log.per, na.rm = T)
# qqplot (life_expectancy_data$Life.expectancy, log.per)
# plot(qqnorm (log.per)); qqline (log.per)
```


# Transformations

``` {r}

life_expectancy_box_cox <- boxcox(life_expectancy_data$Life.expectancy~life_expectancy_data$Status+life_expectancy_data$percentage.expenditure+life_expectancy_data$BMI+life_expectancy_data$Polio+life_expectancy_data$Diphtheria+life_expectancy_data$HIV.AIDS, lambda = seq (-2, 2, 1/10))
life_expectancy_box_cox
life_expectancy_box_cox$x[which.max (life_expectancy_box_cox$y)]

transformed_fit <- lm(life_expectancy_data$Life.expectancy^2~life_expectancy_data$Status+life_expectancy_data$percentage.expenditure+life_expectancy_data$BMI+life_expectancy_data$Polio+life_expectancy_data$Diphtheria+life_expectancy_data$HIV.AIDS)

plot (transformed_fit)
par(mfrow = c(1,2))
qqplot (transformed_fit)
```

## Influence Analysis

```{r}
summary (influence.measures (fit))
dfbetaPlots (fit)
influenceIndexPlot (fit)

summary (influence.measures (transformed_fit))
dfbetaPlots (transformed_fit)
influenceIndexPlot (transformed_fit)

```
















# Response to Feedback

How many data points do you have? Is each country a data point or a combination of country and year a data point? (-1)

Describe members' responsibility on the planned data analysis, not about the proposal. (-1)

Exploratory data analysis described in GOAL is better placed in PLAN.



